---
title: "ISYE 6501 Homework 10"
author: "Mitchell Matsuura"
date: "March 25, 2019"
output:
 word_document: default
pdf_document: default
html_document: default
---
  

###Question 14.1.1
Load in data, do mean, mode and median imputation
```{r}
cancer<-read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data",header = FALSE,stringsAsFactors = FALSE)
str(cancer)
summary(cancer)
View(cancer)
table(cancer$V7)
```
After loading in the data we see that the variable V7 for bare nuclei has 16 observations with missing data labeled as "?".  Observations for this attribute should vary from integers 1-10.

create an index of the missing values and replace '? with 0
```{r}
missing<-which(cancer$V7=='?')
missing

```
index creation is successful since 16 row indicies are returned after calling missing

What % of the data has missing values?  Is it below 5%
```{r}
length(missing)/nrow(cancer)
```
missing values are only a small portion of the data, less than 5%.  So imputation is a reasonable approach to handling missing data.

Is missing data random?  Use prop.table to figure out
```{r}

prop.table(table(cancer$V11))

prop.table(table(cancer[-missing,11]))

prop.table(table(cancer[missing,11]))

```
it apprears that the missing data rows have a different proportion of the response classes than the nonmissing rows. Class 2 is more likely. However we are only missing 16 observations out of 699 so without insider knowledge of the data recording process, we'll just keep this in mind for now.

set V7 to numeric/integer calculate mean, mode, median of nonmissing V7.  Round mean to nearest integer
```{r}
as.integer(cancer$V7)
mean<-round(mean(as.integer(cancer[-missing,7]),na.rm = TRUE), digits = 0)
mean
median<-median(cancer$V7)
median
mode<-as.numeric(names(table(cancer[-missing,7]))[which.max(table(cancer[-missing,7]))])
mode
hist(as.integer(cancer[-missing,7]))
```
mean was rounded to 4.  Median and mode are both equal to 1 since most of our observations are 1.  This is a nice 2 for 1 in our analysis but we'll have to keep in mind the distribution is actually bimodal with 2nd mode at 10, modes are at both ends of the distribution.

replace missing with mean and 
```{r}
cancer[missing,7]=mean
cancer[missing,7]

table(cancer$V7)

as.integer(cancer$V7)
cancer$V11<-as.factor(ifelse(cancer$V11==2,0,1))
```

split the data into build/validate and test with 75% into build/validate and 30% test
```{r}
library(caret)

trainIndex <- createDataPartition(cancer$V11, p = .75, list=FALSE, times = 1)
train<-cancer[trainIndex,]
prop.table(table(train$V11))


cvfolds<-createMultiFolds(train$V11, k=10, times = 5)
control<-trainControl(method = "repeatedcv", number = 10, repeats = 5, 
												 index = cvfolds)
svmgrid<-expand.grid(C=2^c(-6,-5,-4,-3,-2,-1,0,1,2,3,4,5,6,7),sigma=10^-(1:4))
```

build a KNN model with mean imputation
```{r}
set.seed(253)
svmmean<-train(V11~.,
							data = train, 
							method='svmRadial',
							#preProc=c("center","scale"),
							tuneGrid=svmgrid,
							verbose = FALSE,
							trControl=control)
svmmean

set.seed(253)
svmmean2<-train(V11~.,
							data = train, 
							method='svmRadial',
							#preProc=c("center","scale"),
							tuneGrid=svmgrid,
							verbose = FALSE,
							trControl=control,
							metric='Kappa')
svmmean2
```
See if there is a different result if I pick the best model using Kappa metric.  Kappa is a preferred metric for class imbalance.  This data set has a slight imbalance in 0 is twice as likely as 1.  Changing the metric didn't matter since best model is still sigma of 0.01 and C of 1.

Get the confusion matrix on the build set
```{r}
confusionMatrix(predict(svmmean2,newdata=train),train$V11)

```
this is good indication on the build set since we are missclassfying only 5 malignant tumors and 10 benign tumors.

perform mode imputation
```{r}
cancer[missing,7]=mode
cancer[missing,7]
table(cancer$V7)

as.integer(cancer$V7)

train<-cancer[trainIndex,]
```

rerun svmmean2 with mode imputed data and Kappa Metric
```{r}
set.seed(253)
svmmode<-train(V11~.,
							data = train, 
							method='svmRadial',
							#preProc=c("center","scale"),
							tuneGrid=svmgrid,
							verbose = FALSE,
							trControl=control,
							metric='Kappa')
svmmode
```
cross validation selected the same optimal parameter for the svm model with mode imputed missing values of sigma = 0.01  and C=1.  Crossvalidated Kappa is only very slightly lower than mode impuation maybe due to the fact the response distribution is bimodal.  Let's see the confusion matrix on the build set

Get the confusion matrix on the build set
```{r}
confusionMatrix(predict(svmmode,newdata=train),train$V11)

```
no change on the build set confusion matrix since we misclassify 5 malignant and 10 benign like before.  However, I would pay close attention to differences on the test set performance.


