---
title: "ISYE 6501 Homework 4"
author: "Mitchell Matsuura"
date: "February 5, 2019"
output:
  word_document: default
pdf_document: default
html_document: default
---

###Question 7.1
In my industry we often have to make projections of what the cost of paying claims at some point in the future for budgetary and pricing considerations.  As you can imagine it is important to factor in inflation in the products you sell for any business; however, this plays a crucial role for insurance since claims are always paid out after we issue the policy (collect premium) and you don't know exactly when. 

If I were to conduct this time series analysis I would expect the alpha parameter to be nonzero but less than 0.5.  This alpha would indicate that there could be a stronger dependence on past estimates than the actual observations and would correspond to my intuition on the rising cost of medical bills and litigation costs over time in my state.



###Question 7.2
read in text file from class website and examine data
```{r}
temp<-read.table("https://prod-edxapp.edx-cdn.org/assets/courseware/v1/592f3be3e90d2bdfe6a69f62374a1250/asset-v1:GTx+ISYE6501x+1T2019+type@asset+block/temps.txt",header = TRUE)
str(temp)
summary(temp)
View(temp)
```
Convert the data frame to an vector of integers and plot
```{r}
temp_vec<-as.vector(unlist(temp[,2:21]))
head(temp_vec)
tail(temp_vec)
class(temp_vec)
plot(temp_vec)

```

convert data to time series with 123 observations for each of the 20 years of data
```{r}
tempts<-ts(temp_vec, start = 1996, frequency = 123)
View(tempts)
plot(tempts)
```

use decompose in r to see if there is a trend and seasonality
```{r}
plot(decompose(tempts))
```

this graph using decompose (a simple moving average method) indicates a growing intensity in trend around 2010-2013 and a clear seasonality that does not appear to be multiplicative since it is not growing over time.  I'll use decompose with multiplicative time series selected to see if anything happens to trend

```{r}
plot(decompose(tempts, type = "mult"))
```
no real change to trend and observed.  Only change is to the scale of random and seasonal scales.  Let's use the holt winter's method now and see what r estimates for the best values for alpha, beta, gamma

```{r}
temphw<-HoltWinters(tempts, alpha = NULL, beta = NULL, gamma = NULL, seasonal = "multiplicative")
temphw$alpha
temphw$beta
temphw$gamma
temphw$coefficients
temphw$SSE
plot(temphw)
```
From viewing the output we see that the trend parameter is estimated to be very close to zero indicating there is no trend for this period.  However, we have positive nonzero alpha and gamma parameter indicating there is seasonality.

Note SSE is 68904.57

let's try to improve SSE by taking out the unnessary trend parameter beta

```{r}
temphw2<-HoltWinters(tempts, alpha = NULL, beta = NULL, gamma = NULL, seasonal = "additive")
temphw2$alpha
temphw2$beta
temphw2$gamma
temphw2$coefficients
temphw2$SSE
plot(temphw2)
```
Note that switching to additive seasonality the SSE improved from 68904.57 to 66244.25

Try additive seasonality and remove beta

```{r}
temphw3<-HoltWinters(tempts, alpha = NULL, beta = FALSE, gamma = NULL, seasonal = "additive")
temphw3$alpha
temphw3$beta
temphw3$gamma
temphw3$coefficients
temphw3$SSE
plot(temphw3)
```
SSE has not changed after removing beta and alpha & gamma have actually increased slightly after taking out beta so might as well leave beta in so I'll use the previous model with additive seasonality and beta included, temphw2.

Now extract the seasonality component from out temphw2 model
```{r}
temphw2sf<-matrix(temphw2$fitted[,4], nrow = 123)
head(temphw2sf)
```

compute the mean of each day across all 20 years
```{r}
dailyavg<-data.frame(rowMeans(temphw2sf, n = 19))
head(dailyavg)
View(dailyavg)
str(dailyavg)
mean(dailyavg[,1])
library(tidyverse)
ggplot(dailyavg,aes(x=1:123, y= rowMeans.temphw2sf..n...19.))+geom_line()+geom_hline(yintercept = -0.5)+geom_vline(xintercept=78)

```
According to the graph it looks like day 78 is a good base level since it is around the mean of the seasonality means and it seems to be where the avg daily seasonality is head back to this level at the end of the graph.  I will  use this base level to see how many days later summer is detected using CUSUM.

First I need to find the sd of dailyavg resembling true summer weather, averages for July-Aug, and apply formulas for C and T using 0.5 x standard deviation and 5 x standard deviation and revise if necessary
```{r}
std<- sd(dailyavg$rowMeans.temphw2sf..n...19[1:62])#standard deviation for July-Aug
C<- std * 0.5
T<- std * 5 
C
T

```

add CUSUM columns for dailyavg dataframe with C = 1.616309 and T = 16.16309
```{r}
s = list()
  dailyavg$s= 0
  CUSUM<-function(C,T,cl){
for (j in cl){
  for (i in 1:nrow(dailyavg)){
  #s[i-1,j] = 0
if (i-1 != 0){
  
  s <- max(0,dailyavg$s[i-1]+(mean(dailyavg[,j])-dailyavg[i,j]-C))#CUSUM to detect decrease
}
    else{ s<- 0}
    if(s>=T){s}
    else{s<-0}
    
dailyavg$s[[i]]<-s
  }
  }
  return(data.frame(1:123,dailyavg[,j],dailyavg$s)[78:123,])}#limit the output
  CUSUM(C=1.616309,T=16.16309,cl=1)
```
Using CUSUM and selected levels of C and T, I detected no change to the end of summer.  This is a consistent conclusion to HW3 where we didn't find consistent changes to indicate if summer is ending later or earlier.

Try running daily avg as a time series with the exponential smoothing (estimate only alpha) instead of CUSUM since they are used to do the same thing, making a weighted avg estimates based on current observation and past projections.

convert dailyavg to ts format
```{r}
seas_vec<-as.vector(unlist(dailyavg[,1]))
class(seas_vec)
plot(seas_vec)

seasts<-ts(seas_vec)
View(seasts)
plot(seasts)
```

Run exponential smoothing on avg seasonal time series
```{r}
seases<-HoltWinters(seasts, alpha = NULL, beta = FALSE, gamma = FALSE, seasonal = "additive")
seases$x
seases$alpha
seases$coefficients
seases$SSE
plot(seases)
```

Here, we want to focus on the alpha parameter, 0.999956.  The expontial smoothing found optimal alpha to be very close to 1 or in other words, we are better off relying on actual observations than our prior estimates of the mean.  Therefore we can conclude the same as CUSUM that there is no systematic changes overtime to indicate summer is ending later over the 20 year period.  Large changes are due to random variation in the data.


